### 2018-01-22 conclusion.tex
initial upload this is a short version of my bachelor thesis. had to do it in conjunction with the 'ringvorlesung' in master.

\section{Conclusion and Future Work}
A correlation based approach to the 6DOF image registration problem consisting of 3D translation and rotation was introduced.
The algorithm is capable of determining arbitrary rotations of an object between two voxel data sets.
It leverages the Fourier transform to decouple the rotational from translational information.
The rotation is calculated using sphere surfaces with different rotation axes of the data sets Fourier magnitude spectra, where the sphere surface with highest cross-correlation coefficient indicates the correct axis and the location of this coefficient indicates the angle of rotation around that axis.

The accuracy of the algorithm is dependent on the resolution of the data sets but could be improved when using a sub pixel accurate cross-correlation as proposed by Campbell and Wu~\cite{campbell08gradientcrosscorr}.

For high resolution data sets, a single sphere surface may contain insufficient data due to the small frequency range of a single radius.
More information could be obtained when integrating over a radius range in sphere surface extraction. 
B\"ulow and Birk~\cite{bulow13dof6registration} provided evidence for accuracy improvement in their approach when using multiple layers of several radii.