### 2018-07-05 seminar_vof/paper/paper.tex
writing variational flow recap

\documentclass[journal]{vgtc} 
\usepackage{hs-vis_ss10}

%% Please note that the use of figures other than the optional teaser
%% is not permitted on the first page of the journal version.  Figures
%% should begin on the second page and be in CMYK or Grey scale
%% format, otherwise, colour shifting may occur during the printing
%% process.  Papers submitted with figures other than the optional
%% teaser on the first page will be refused.

%% These three lines bring in essential packages: ``mathptmx'' for
%% Type 1 typefaces, ``graphicx'' for inclusion of EPS figures. and
%% ``times'' for proper handling of the times font family.

\usepackage{mathptmx} 
\usepackage{graphicx}
\usepackage{times}
%\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{bigints}
\usepackage{xfrac}
\usepackage{bm}

\newcommand{\lam}{\lambda}
\newcommand{\trans}{\quad\bigg|\;}
\newcommand{\Trans}{\quad\Bigg|\;}
% matrix shortcut
\newcommand{\mat}[1]{\begin{pmatrix} #1 \end{pmatrix}}
% matrix shortcut with increased vertical spacing
\newcommand{\Mat}[1]{\begingroup
\renewcommand*{\arraystretch}{1.5}
\mat{#1}
\endgroup}
% matrix format but without braces (secretly a matrix)
\newcommand{\secretMat}[2][r]{\begin{matrix*}[#1] #2 \end{matrix*}}
% matrix format but without braces (secretly a matrix) with increased vertical spacing
\newcommand{\SecretMat}[2][r]{\begingroup
\renewcommand*{\arraystretch}{1.7}
\secretMat[#1]{#2}
\endgroup}
% shortcuts for df/dx stuff in derivative
\newcommand{\der}{\partial}
\newcommand{\deriv}[2]{\frac{\der #1}{\der #2}}
% make star '*' become \cdot (normal multiplication sign)
\mathcode`\*="8000
{\catcode`\*\active\gdef*{\cdot}}
\newcommand{\Real}{\mathbb{R}}
\newcommand{\flow}{\vec{u}}
\newcommand{\x}{\vec{x}}
\newcommand{\y}{\vec{y}}
\newcommand{\ev}{\vec{r}}
\newcommand{\aaux}{\vec{a}}
\newcommand{\baux}{\vec{b}}
\newcommand{\T}{^\top}
\newcommand{\argmin}{\mathop{\mathrm{argmin}}}

%% allow for this line if you want the electronic option to work
%% properly
\vgtcinsertpkg

%% author name
\author{Roy G. Biv}

%% paper title
\title{Visualisierung von Regenbogenfarben}

%% short title for header
\shorttitle{Visualisierung von Regenbogenfarben}

%% Abstract section.
\abstract{%
  Hier sollte eine kurze Zusammenfassung der Ausarbeitung stehen. Der
  Umfang sollte 150 bis 250 Worte betragen. Duis autem vel eum iriure
  dolor in hendrerit in vulputate velit esse molestie consequat, vel
  illum dolore eu feugiat nulla facilisis at vero eros et accumsan et
  iusto odio dignissim qui blandit praesent luptatum zzril delenit
  augue duis dolore te feugait nulla facilisi. Lorem ipsum dolor sit
  amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod
  tincidunt ut laoreet dolore magna aliquam erat volutpat.

  Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
  suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis
  autem vel eum iriure dolor in hendrerit in vulputate velit esse
  molestie consequat, vel illum dolore eu feugiat nulla facilisis at
  vero eros et accumsan et iusto odio dignissim qui blandit praesent
  luptatum zzril delenit augue duis dolore te feugait nulla facilisi.
} % end of abstract

%% Uncomment below to include a (optional) teaser figure.
%\teaser{ \centering
%  \includegraphics[width=16cm]{images/CypressView}
%  \caption{In den Wolken: Vancouver von Cypress Mountain. Auf der
%    ersten Seite d"urfen keine Grafiken au"ser dieser optionalen
%    Aufmachgrafik (Teaser) abgebildet sein.}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title
%%   block.

%%   the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

Die Ausarbeitung darf eine L"ange von 6 Seiten nicht unterschreiten
und 8 Seiten nicht "uberschreiten. Dies beinhaltet Text, Abbildungen
und Referenzen. Bei Ausarbeitungen mit mehr als 8 Seiten werden nur
die ersten 8 Seiten zur Bewertung herangezogen. F"ur die Form der
Ausarbeitung steht auf der Seminarwebseite diese \LaTeX{}-Vorlage
bereit. Diese Vorlage ist bindend und darf nicht abge"andert
werden. Abgaben in anderer Form werden nicht akzeptiert.

\section{Instruktionen}

Bitte nachfolgende Abschnitte sorgf"altig durchlesen.

\subsection{Sprache}

Die Ausarbeitung kann entweder auf Deutsch oder Englisch geschrieben
werden.

\subsection{Abbildungen und Tabellen}

Alle Abbildungen (siehe Abb.\ \ref{fig:sample}) und Tabellen (Tabelle\
\ref{tab:vis_accept}) sollten zentriert sein
(\verb|\centering|). Abbildungen "uber beide Textspalten
(Abb. \ref{fig:multicolumn}) k"onnen mit
\verb|\begin{figure*}|\ldots\verb|\end{figure*}| eingef"ugt werden.

\subsection{Referenzen}

Literaturangaben wie beispielsweise Levoy \cite{levoy:1989:DSV} werden
mit Hilfe von BibTeX erzeugt. Dazu werden die Referenzen in die
Literaturliste (hier \emph{literatur.bib}) eingetragen und
entsprechend mit \verb|\cite| referenziert.

\subsection{\LaTeX-"Ubersetzung}

Die \LaTeX-Datei kann mit \emph{latex} oder \emph{pdflatex} "ubersetzt
werden. Dabei ist zu beachten, dass f"ur die "Ubersetzung mit
\emph{latex} die Grafiken in Postscript (eps) vorliegen, f"ur
\emph{pdflatex} entsprechend als jpg, png oder pdf.  Der Ablauf ist
dabei der folgende:
\begin{enumerate}
\item \verb|pdflatex <quelldatei.tex>|
\item \verb|bibtex <quelldatei>|
\item \verb|pdflatex <quelldatei.tex>| (evtl. mehrfach)
\end{enumerate}
Alternativ kann auch das mitgelieferte Makefile verwendet werden.

\section{Introduction}

In computer vision, the field of optic flow deals with the estimation of a 2-dimensional flow field based on two consecutive frames of an image sequence.
\\(mehr dazu warum optic flow geil ist)\\
Different methods for flow estimation exist, such as ... ... and variational optic flow which is based on the calculus of variations.
In the following the concept of variational flow estimation is recapitulated and the approach by \verb|\cite{daspaper}| with order adaptive regulalarization is presented.

\section{Variational Optic Flow}
In this section the general framework for flow estimation using variational methods is recapitulated.
\\\\
Given two subsequent images
$I_0,I_1 ~:~ \Omega \subset \Real^2 \to \Real$ \\
of an image sequence, we seek the displacement vector field 
\\$\flow = (u~v)\T ~:~ \Omega \subset \Real^2 \to \Real^2$ 
that maps 
$I_0$ onto $I_1$ 
such that\\
$I_0(\x) = I_1(\x+\flow(\x))$.
\\\\
To estimate the flow function $\flow$ an energy functional is minimized. 
Typically this functional consists of two terms, a data term $D(\flow)$ and a regularization term $R(\flow)$.
\begin{align*}
\flow' &= \argmin_{\flow} E(\flow)
\\
E(\flow) &= D(\flow) + \alpha R(\flow)
\end{align*}
In this framework, the data term is used to penalize flow functions which violate certain constancy assumptions on image features, such as the brightness constancy assumtion 
$D_{brght}(\flow) = \int_\Omega (I_0(\x)-I_1(\x+\flow(\x))^2\;d\x$.
As the data term is insufficient for solving this problem and would yield multiple ambiguous solutions as illustrated in \ref{fig:ambigdataterm}, the regularization term is used to overcome this inconvenience and penalize flow functions that are unlikely.
This is done by setting up constraints for the flow, such as the first order smoothness assumption
$R_{smooth}(\flow) = \int_\Omega || \nabla u ||_2^2 + || \nabla v ||_2^2\;d\x$.
This assumption enforces constant flow due to penalizing non vanishing gradients, which amount to changes in the vector field.

\begin{figure}[tb]
  \centering
  \includegraphics[width=1.5in]{images/ambiguousflow.png}
  \caption{\label{fig:ambigdataterm} Beispielillustration.}
\end{figure}

\section{Exposition}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum~\cite{ware:2004:IVP}. Stet clita kasd gubergren, no sea takimata
sanctus est Lorem ipsum dolor sit amet.

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat~\cite{kindlmann:1999:SAG}.

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo
consequat~\cite{levoy:1989:DSV}. Duis autem vel eum iriure dolor in
hendrerit in vulputate velit esse molestie consequat, vel illum dolore
eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua.

\begin{figure}[tb]
  \centering
  \includegraphics[width=1.5in]{sample}
  \caption{\label{fig:sample} Beispielillustration.}
\end{figure}

\begin{equation}
  \sum_{j=1}^{z} j = \frac{z(z+1)}{2}
\end{equation}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

\begin{table}
  %% Table captions on top in journal version
  \caption{\label{tab:vis_accept} Vis Paper Acceptance Rate}
  \scriptsize
  \begin{center}
    \begin{tabular}{cccc}
      Year & Submitted & Accepted & Accepted (\%)\\
      \hline
      1994 &  91 & 41 & 45.1\\
      1995 & 102 & 41 & 40.2\\
      1996 & 101 & 43 & 42.6\\
      1997 & 117 & 44 & 37.6\\
      1998 & 118 & 50 & 42.4\\
      1999 & 129 & 47 & 36.4\\
      2000 & 151 & 52 & 34.4\\
      2001 & 152 & 51 & 33.6\\
      2002 & 172 & 58 & 33.7\\
      2003 & 192 & 63 & 32.8\\
      2004 & 167 & 46 & 27.6\\
      2005 & 268 & 88 & 32.8\\
      2006 & 228 & 63 & 27.6
    \end{tabular}
  \end{center}
\end{table}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

\subsection{Mezcal Head}

Duis autem~\cite{Lorensen:1987:MCA} vel eum iriure dolor in hendrerit
in vulputate velit esse molestie consequat, vel illum dolore eu
feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt
ut laoreet dolore magna aliquam erat volutpat%
\footnote{Fu"snoten erscheinen an der Unterkante der Spalte. Sie
  sollten jedoch vermieden werden, da der Lesefluss gest"ort wird.}.

\subsubsection{Ejector Seat Reservation}

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo
consequat~\cite{Nielson:1991:TAD}. Duis autem vel eum iriure dolor in
hendrerit in vulputate velit esse molestie consequat, vel illum dolore
eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi.

\paragraph{Rejected Ejector Seat Reservation}

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie

\section{Conclusion}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum.

\bibliographystyle{abbrv}
%% use following if all content of bibtex file should be shown
% \nocite{*}
\bibliography{literatur}
\end{document}