### 2018-07-28 mlr_summary.tex
started mlrsummary

% !TeX spellcheck = en_GB
\documentclass[12pt]{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{xfrac}
\usepackage{esvect}
\setlength{\parindent}{0pt}
\setlength{\parskip}{.36cm plus0.09cm minus0.09cm}
\usepackage[left=2.25cm, right=2.25cm, top=2.25cm, bottom=2.25cm]{geometry}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{esvect}
\usepackage{csquotes}

\newcommand{\lam}{\lambda}
\newcommand{\trans}{\quad\bigg|\;}
\newcommand{\Trans}{\quad\Bigg|\;}
% matrix shortcut
\newcommand{\mat}[1]{\begin{pmatrix} #1 \end{pmatrix}}
% matrix shortcut with increased vertical spacing
\newcommand{\Mat}[1]{\begingroup
\renewcommand*{\arraystretch}{1.5}
\mat{#1}
\endgroup}
% matrix format but without braces (secretly a matrix)
\newcommand{\secretMat}[2][r]{\begin{matrix*}[#1] #2 \end{matrix*}}
% matrix format but without braces (secretly a matrix) with increased vertical spacing
\newcommand{\SecretMat}[2][r]{\begingroup
\renewcommand*{\arraystretch}{1.7}
\secretMat[#1]{#2}
\endgroup}
% shortcuts for df/dx stuff in derivative
\newcommand{\der}{\partial}
\newcommand{\deriv}[2]{\frac{\der #1}{\der #2}}
\newcommand{\eqns}[1]{\begin{flalign} #1 \end{flalign}}
\newcommand{\eqnsnn}[1]{\begin{flalign*} #1 \end{flalign*}}
\newcommand{\argmin}{\mathop{\mathrm{argmin}}}
\newcommand{\dom}[1]{\mathbb{#1}}
\newcommand{\T}{^\top}
\newcommand{\equivalent}{\Leftrightarrow}
% make star '*' become \cdot (normal multiplication sign)
\mathcode`\*="8000
{\catcode`\*\active\gdef*{\cdot}}

\begin{document}
  \begin{center}
    \Large\textbf{Summary - Machine Learning}\\
    \vspace{0.25cm}
    \large{David H\"agele}, \normalsize{\today}
   \end{center}

\section{Machine Learning Briefly}
The goal of machine learning is to estimate or \enquote{learn} a mapping $f$ from one domain $\dom{X}$ to another $\dom{Y}$ by using a set of data points.
\eqnsnn{
f : \dom{X} \to \dom{Y} &&
}
In the field of supervised learning a data set $D$ usually contains observed data points of domain $\dom{X}$ and corresponding points of domain $\dom{Y}$. 
The objective is to find a mapping $f_\text{optimal}$ that minimizes an error $err$ between observed and estimated $y$.
\eqnsnn{&
D = \{ x_i , y_i \}\quad|\quad x_i\in\dom{X},~ y_i \in \dom{Y},~ i\in\dom{N}
&&\\&
err : \dom{Y}\times\dom{Y} \to \dom{R}
&&\\&
f_\text{optimal} = \argmin_f\;err(y,f(y))
&&}
In the field of unsupervised learning there are no corresponding observed points of domain $\dom{Y}$.
Optimality is not measured using an error to the observation, instead other objectives are to be fulfilled.

\section{Linear Regression}
One of the most basic algorithms for estimating a real valued function is linear regression.
Given a set of multivariate inputs $x_i \in \dom{R}^m$ and corresponding real valued outputs $y_i \in \dom{R}$ the following \enquote{loss} $L$ is to be minimized by a linear function $f_\beta:\dom{R}^m\to\dom{R}$.
\eqns{&
L(\beta) = \sum_{i=1}^n (\;y_i - f_\beta(x_i)\;)^2 
&&\\&
f_\beta(x) = x\T\beta \quad\text{with}\quad\beta\in\dom{R}^m
&&}
Lets call the matrix of transposed $x_i$'s $X$ and the vector of $y_i$'s $Y$.
This allows for rewriting the loss as follows.
\eqnsnn{&
L(\beta) = (Y-X\beta)\T(Y-X\beta)
&&\\&
X = \mat{x_1\T\\\vdots\\x_n\T} \quad Y = \mat{y_1\\\vdots\\y_n}
&&}
When minimizing we seek the roots of the derivative of the loss, by doing so we can derive the optimal solution for beta.
\eqnsnn{
\deriv{L(\beta)}{\beta} = 0 ~&\equivalent~ (Y-X\beta)\T X + (Y-X\beta)\T X = 2*(Y-X\beta)\T X = 0
&&\\&
\equivalent (Y-X\beta)\T X = X\T Y - X\T X\beta = 0
&&\\&
\equivalent X\T Y = X\T X\beta
&&\\&
\equivalent (X\T X)^{-1}X\T Y = \beta
&&
}
\subsection{Non-Linear Features}
Sadly the very easy linear regression algorithm only allows for linear functions to be estimated. 
Fortunately we can make things non-linear by defining a feature function which maps an input $x$ to feature space like so:
\eqnsnn{
\phi(x) \in \dom{R}^l \quad\text{with}\quad l\in\dom{N}
&&}
For real valued $x\in\dom{R}$ we can create the quadratic feature vector using the following.
\eqnsnn{
\phi(x) = \mat{1&x&x^2}\T
&&}
When using linear regression on these features we end up with a function that is quadratic in $x$ but linear in features $\phi(x)$.
\eqnsnn{&
X = \mat{\phi(x_1)\T\\\vdots\\\phi(x_n)\T}
&&\\&
f_\beta(x) = \phi(x)\T\beta = \phi(x)_1*\beta_1 + \phi(x)_2*\beta_2 + \phi(x)_3*\beta_3 = \beta_1 + x\beta_2 + x^2\beta_3
&&
}

\subsection{Regularization}
Maybe you may have noticed that the linear regression algorithm may fail in case that $(X\T X)$ is not invertible.
To ensure invertibility a tiny quantity can be added to the diagonal of the matrix $(X\T X + \lambda I)$.
Interestingly, the optimal solution to the following loss has this exact form.
\eqnsnn{&
L(\beta) = (Y-X\beta)\T(Y-X\beta) + \lambda*\beta\T\beta
&&\\&
\deriv{L(\beta)}{\beta} = 0 \equivalent 2(Y-X\beta)X\T + 2\lambda*\beta\T = 0
\equivalent (Y-X\beta)X\T + \lambda*\beta\T = 0
&&}

\end{document}