### 2018-07-28 mlr_summary.tex
linreg done, also done features,regularization and cross validation

% !TeX spellcheck = en_GB
\documentclass[12pt]{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{xfrac}
\usepackage{esvect}
\setlength{\parindent}{0pt}
\setlength{\parskip}{.36cm plus0.09cm minus0.09cm}
\usepackage[left=2.25cm, right=2.25cm, top=2.25cm, bottom=2.25cm]{geometry}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{esvect}
\usepackage{csquotes}
\usepackage{algpseudocode}

\newcommand{\lam}{\lambda}
\newcommand{\trans}{\quad\bigg|\;}
\newcommand{\Trans}{\quad\Bigg|\;}
% matrix shortcut
\newcommand{\mat}[1]{\begin{pmatrix} #1 \end{pmatrix}}
% matrix shortcut with increased vertical spacing
\newcommand{\Mat}[1]{\begingroup
\renewcommand*{\arraystretch}{1.5}
\mat{#1}
\endgroup}
% matrix format but without braces (secretly a matrix)
\newcommand{\secretMat}[2][r]{\begin{matrix*}[#1] #2 \end{matrix*}}
% matrix format but without braces (secretly a matrix) with increased vertical spacing
\newcommand{\SecretMat}[2][r]{\begingroup
\renewcommand*{\arraystretch}{1.7}
\secretMat[#1]{#2}
\endgroup}
% shortcuts for df/dx stuff in derivative
\newcommand{\der}{\partial}
\newcommand{\deriv}[2]{\frac{\der #1}{\der #2}}
\newcommand{\eqns}[1]{\begin{flalign} #1 \end{flalign}}
\newcommand{\eqnsnn}[1]{\begin{flalign*} #1 \end{flalign*}}
\newcommand{\argmin}{\mathop{\mathrm{argmin}}}
\newcommand{\dom}[1]{\mathbb{#1}}
\newcommand{\T}{^\top}
\newcommand{\equivalent}{\Leftrightarrow}
% make star '*' become \cdot (normal multiplication sign)
\mathcode`\*="8000
{\catcode`\*\active\gdef*{\cdot}}

\begin{document}
  \begin{center}
    \Large\textbf{Summary - Machine Learning}\\
    \vspace{0.25cm}
    \large{David H\"agele}, \normalsize{\today}
   \end{center}

\section{Machine Learning Briefly}
The goal of machine learning is to estimate or \enquote{learn} a mapping $f$ from one domain $\dom{X}$ to another $\dom{Y}$ by using a set of data points.
\eqnsnn{
f : \dom{X} \to \dom{Y} &&
}
In the field of supervised learning a data set $D$ usually contains observed data points of domain $\dom{X}$ and corresponding points of domain $\dom{Y}$. 
The objective is to find a mapping $f_\text{optimal}$ that minimizes an error $err$ between observed and estimated $y$.
\eqnsnn{&
D = \{ x_i , y_i \}\quad|\quad x_i\in\dom{X},~ y_i \in \dom{Y},~ i\in\dom{N}
&&\\&
err : \dom{Y}\times\dom{Y} \to \dom{R}
&&\\&
f_\text{optimal} = \argmin_f\;err(y,f(y))
&&}
In the field of unsupervised learning there are no corresponding observed points of domain $\dom{Y}$.
Optimality is not measured using an error to the observation, instead other objectives are to be fulfilled.

\section{Linear Regression}
One of the most basic algorithms for estimating a real valued function is linear regression.
Given a set of multivariate inputs $x_i \in \dom{R}^m$ and corresponding real valued outputs $y_i \in \dom{R}$ the following \enquote{loss} $L$ is to be minimized by a linear function $f_\beta:\dom{R}^m\to\dom{R}$.
\eqns{&
L(\beta) = \sum_{i=1}^n (\;y_i - f_\beta(x_i)\;)^2 
&&\\&
f_\beta(x) = x\T\beta \quad\text{with}\quad\beta\in\dom{R}^m
&&}
This is the mean squared error and we want to minimize it using a linear function.
Lets call the matrix of transposed $x_i$'s $X$ and the vector of $y_i$'s $y$.
This allows for rewriting the loss as follows.
\eqnsnn{&
L(\beta) = (y-X\beta)\T(y-X\beta)
&&\\&
X = \mat{x_1\T\\\vdots\\x_n\T} \quad y = \mat{y_1\\\vdots\\y_n}
&&}
When minimizing we seek the roots of the derivative of the loss, by doing so we can derive the optimal solution for beta.
\eqnsnn{
\deriv{L(\beta)}{\beta} = 0 ~&\equivalent~ (y-X\beta)\T X + (y-X\beta)\T X = 2*(y-X\beta)\T X = 0
&&\\&
\equivalent (y-X\beta)\T X = X\T y - X\T X\beta = 0
&&\\&
\equivalent X\T y = X\T X\beta
&&\\&
\equivalent (X\T X)^{-1}X\T y = \beta
&&
}
\subsection{Non-Linear Features}
Sadly the very simple linear regression algorithm only allows for linear functions to be estimated. 
Fortunately we can make things non-linear by defining a feature function which maps an input $x$ to feature space like so:
\eqnsnn{
\phi(x) \in \dom{R}^l \quad\text{with}\quad l\in\dom{N}
&&}
For real valued $x\in\dom{R}$ we can create the quadratic feature vector using the following.
\eqnsnn{
\phi(x) = \mat{1&x&x^2}\T
&&}
When using linear regression on these features we end up with a function that is quadratic in $x$ but linear in features $\phi(x)$.
\eqnsnn{&
X = \mat{\phi(x_1)\T\\\vdots\\\phi(x_n)\T}
&&\\&
f_\beta(x) = \phi(x)\T\beta = \phi(x)_1*\beta_1 + \phi(x)_2*\beta_2 + \phi(x)_3*\beta_3 = \beta_1 + x\beta_2 + x^2\beta_3
&&
}
Even more sophisticated features could be using gaussian functions at the datapoints which then amounts to a radial basis function.
\eqnsnn{&
\phi(x) = \mat{1 & \phi_1(x) & \cdots & \phi_N(x)}\T \quad\text{with}\quad \phi_i(x) = e^{-\frac{1}{2}||x_i-x||_2^2}
&&}

\subsection{Regularization}
Maybe you may have noticed that the linear regression algorithm may fail in case that $(X\T X)$ is not invertible.
To ensure invertibility a tiny quantity can be added to or subtracted from the diagonal of the matrix like so: $(X\T X - \lambda I)$.
Interestingly, the optimal solution to the following loss has this exact form.
\eqnsnn{
L(\beta) = (y&-X\beta)\T(y-X\beta) + \lambda*\beta\T\beta
&&\\
\deriv{L(\beta)}{\beta} = 0 &\equivalent 2(y-X\beta)\T X + 2\lambda*\beta\T = 0
\equivalent (y-X\beta)\T X + \lambda*\beta\T = 0
&&\\
&\equivalent X\T y - X\T X\beta + \lambda*\beta = 0 
\equivalent X\T y = X\T X\beta - \lambda*\beta = (X\T X - \lambda I) \beta
&&\\
&\equivalent (X\T X - \lambda I)^{-1}X\T y = \beta
&&} 
The loss function can be viewed as a combination of penalties for $\beta$.
The first penalty is the mean squared error from before which is large when $\beta$ cannot reproduce the samples $y$ well.
The second penalty is for $\beta$ of long lenghts, which means that short $\beta$ with small coefficients are preferred.
These two penalties probably are contradictive and the optimal beta will be somewhere in between the individual minima for the penalties.
This kind of regularization using $\lambda\beta\T\beta$ is called $L^2$ regularization and is a special case of the general Tikhonov regularization formulating constraints on $\beta$ as $\beta\T\Gamma\T\Gamma\beta$ with Tikhonov matrix $\Gamma$.
$\Gamma\beta$ is a linear mapping which can be chosen to also formulate special constraints on beta like differences between coefficients, or projections to different vector spaces.

When we use only half of the data set for \enquote{training} our model i.e. solving for $\beta$ we can check how well it reproduces the other half of the data set, which indicates how well the model generalizes.
While the model may be very accurate for the data used for training, it may be inaccurate for the data used for \enquote{validation} (the other half), which means that the model is \enquote{over fitted} to the training data and does not approximate the actually underlying function well.
In our case we can use regularization to allow the model to have a less perfect fit to the data when it instead fulfills the other objective of being small.
The degree to which the model can violate the mean squared error depends on $\lambda$.

\subsection{Cross Validation}
In order to find the perfect regularization parameter $\lambda$ in order for our model to generalize well we need to try some out.
This can be done using cross validation.
In cross validation, we divide our data set into $K$ euqually sized parts (randomly), then we train our model using K-1 parts of the data and validate the model by computing the mean suared error using the missing part.
This is done k times so that each part has been used for validation once and the errors are averaged.
The following pseudo code should make the algorithm clear.
\begin{algorithmic}[1]
\State Split data $D$ into $K$ equally sized parts $\{D_1,\dots,D_K\}$ randomly
\For{$k=1\dots K$}
\State train on data $D_{\backslash D_k}$ yielding $\beta_i$
\State calculate MSE using $D_k$ ~ $err_k = L^{MSE}(\beta_k,D_k)$
\EndFor
\State calculate total MSE as $err=\frac{1}{K}\sum_{k=1}^K \frac{err_k}{|D_k|}$ 
\end{algorithmic}
The total error will be a measure for generalization of the model, using this measure an optimal $\lambda$ can be found by trying different $\lambda$ and adaptively tightening the search space.
For the start it may be useful to try lambdas of different orders of magnitude first.

\end{document}